Neural networks for solving regression and classification problems

### 2022-11-15 doc/paper.tex
poetry, shakespeare even

From observing FIG CONSTANT V HESSIAN we see that Newton's method using Hessian inversion greatly outperforms the constant scheduler by reaching an MSE of 0.02 within the first epoch. This is due to the Hessian inversion in the first epoch in Newton's method being equivalent to the analytical solution.  By comparison, the constant scheduler given 200 epochs reaches an MSE of 0.21. Given this result, it would seem indefensible to claim that Newtons method requires improvements, especially considering the time it took the Newton method 0.1512 seconds to compute 200 epochs, again beating the result from the constant scheduler which took 0.4991 seconds \footnote{this result should be taken with a grain of salt, given that the constant scheduler is run as a part of a larger architecture, whilst the Hessian calculation is done in a seperate script. Still, this provides an intuition that Newton's method does not take much time to compute this specific problem}. This result is due to the Hessian matrix only being a mere 55x55 matrix for this problem, such that the matrix inversion will not be computationally complex. However, if we were to increase the complexity of the problem, Newtons method would rapidly increase in computational complexity as it is a large matrix inversion with many floating point operations. For problems of such a small scale, when our features are based on a polynomial approximation Franke Function, it can be important to remember that we don't necessarily have to turn to these fancier solutions as the simple Newton method will do the job perfectly. Keeping this lesson in mind, we wish to further investigate the various scheduler methods in order to gain a better understanding of how well they perform, which will become useful for larger problems.

@@

By introducing mini-batch gradient descent of different batch sizes, we can observe from BATCH SIZE FIG that the scheduler methods perform better when optimized for different batch sizes. Both the constant and the momentum scheduler seem to vary equally over batch sizes, with the optimal batch size being equal to one. This means that our schedulers work the best for a fully stochastic gradient descent. The MSE with full gradient descent, as can be seen in BATCH SIZE FIG when batch size equals 320, provides an MSE of 
A full comparison of batch sizes for all the different schedulers can be viewed in APPENDIX in the appendix.

Furthermore, with the introduction of various adaptive scheduler methods such as RMS prop, adagrad, adagrad with momentum and adam we can observe from FIGURE ALL that we can reach even greater results in fewer epochs. We can also observe that the different schedulers all reach an MSE of 0.02 within a couple of epochs, but in order to get a better understanding of the computational cost we have also compared the time it took to fit the data to an MSE of 0.2. Looking at TIME TABLE FOR ALL SCHEDULERS, Adagrad took the least time, beating out our baseline constant scheduler by 0,222106 seconds. Based on FIG ALL, Adam seems to reach an MSE of 0.2 the fastest when strictly looking at the number of epochs required, however, whilst Adagrad requires more epochs the computations are computed faster in our implementation. Surprisingly, RMS prop performed more poorly than expected compared to other adaptive schedulers.

@@

Another observation from FIG NUM  is that the schedulers final error for100 epochs are more or less the same. As the largest change in MSE occurs in the first 50 epochs as seen in ALL FIGS, perhaps the other schedulers are given enough time to catch up as the best performing schedulers, such as Adagrad, starts slowing down the rate at which the MSE is decreased. However, as can be seen in HEATMAP FIG, Adagrad offers a 12\% \footnote{$\frac{constant MSE}{adagrad MSE} * 100 = \frac{0.018}{0.015} * 100 = 12\%$} decrease in MSE compared to the constant scheduler. It is clear from the results thus far that Adagrad is the best scheduler for approximating the Franke function when using linear regression with gradient descent. We believe that the reason Adagrad wins out is that these result are based on training for 100 epochs. The downfall of Adagrad in the literature seems to be that the aggressive decaying of the learning rate hinders a quick convergence deeper into training SOURCE https://towardsdatascience.com/learning-parameters-part-5-65a2f3583f7d. We are thus cautionary to claim that Adagrad will do the best job over a much larger amount of epochs when compared to adaptive schedulers that claim to improve this aspect.

To test this, we compared Adagrad with Adam over 5000 epochs. Both schedulers performed equally, both resulting in an MSE of 0.017 with Adagrad taking the edge with an MSE 0.0002 lower. Scikit MSE 0.005.

>>>>

FIG CONSTANT V HESSIAN

From observing FIG CONSTANT V HESSIAN we see that Newton's method using Hessian inversion greatly outperforms the constant scheduler by reaching an MSE of 0.02 for the testing data within the first epoch. This is due to the Hessian inversion in the first epoch in Newton's method being equivalent to the analytical solution.  By comparison, the constant scheduler given 200 epochs reaches an MSE of 0.198 for the testing data. Given this result, it would seem indefensible to claim that Newtons method requires improvements, especially considering the time it took the Newton method 0.1512 seconds to compute 200 epochs, again beating the result from the constant scheduler which took 0.4991 seconds \footnote{this result should be taken with a grain of salt, given that the constant scheduler is run as a part of a larger architecture, whilst the Hessian calculation is done in a seperate script. Still, this provides an intuition that Newton's method does not take much time to compute this specific problem}. This result is due to the Hessian matrix only being a mere 55x55 matrix for this problem, such that the matrix inversion will not be computationally complex. However, if we were to increase the complexity of the problem, Newtons method would rapidly increase in computational complexity as it is a large matrix inversion with many floating point operations. For problems of such a small scale, when our features are based on a polynomial approximation Franke Function, it can be important to remember that we don't necessarily have to turn to these fancier solutions as the simple Newton method will do the job perfectly. Keeping this lesson in mind, we wish to further investigate the various scheduler methods in order to gain a better understanding of how well they perform, which will become useful for larger problems.

@@

BATCH SIZE FIG (should mention in text that 320 is the size of the training data. Should mention base batch size for grid search was 18)

By introducing mini-batch gradient descent of different batch sizes, we can observe from BATCH SIZE FIG that the scheduler methods perform better when optimized for different batch sizes over 100 epochs. Both the constant and the momentum scheduler seem to vary equally over batch sizes, with the optimal batch size being equal to one. This means that our schedulers work the best for a fully stochastic gradient descent (SGD). The MSE with full gradient descent, as can be seen in BATCH SIZE FIG when batch size equals 320, provides an MSE of 4.211 of the test data, whilst SGD provides an MSE of 0.28 for the test data in an equal amount of time. Because we first gridsearched the optimal values of eta and lambda first with a predefined base batch size, we must acknowledge that BATCH SIZE FIG will have some bias towards the batch size used when found the optimal hyperparameters. However, given that we have gridsearched with different base batch sizes accross our testing, we have found the results to be similar accross different base batch sizes, and that this bias which we induced to save computational time was insignificant. A full comparison of MSE over batch sizes for all the different schedulers can be viewed in APPENDIX in the appendix.

FIG ALL
TIME TABLE FOR ALL SCHEDULERS

Furthermore, with the introduction of various adaptive scheduler methods such as RMS prop, adagrad, adagrad with momentum and adam we can observe from FIGURE ALL that we can reach even greater results in fewer epochs. We can also observe that the different schedulers all reach an MSE of 0.02 within a couple of epochs, but in order to get a better understanding of the computational cost we have also compared the time it took to fit the data to an MSE of 0.2 of the test data. Looking at TIME TABLE FOR ALL SCHEDULERS, Adagrad took the least time, beating out our baseline constant scheduler by 0,222106 seconds. Based on FIG ALL, Adam seems to reach an MSE of 0.2 the fastest when strictly looking at the number of epochs required, however, whilst Adagrad requires more epochs the computations are computed faster in our implementation. Surprisingly, RMS prop performed more poorly than expected compared to other adaptive schedulers.

FIGURE NUM (heatmap)

@@

Another observation from FIG NUM is that the schedulers final error for 100 epochs are more or less the same. As the largest change in MSE occurs in the first 50 epochs as seen in ALL FIGS, perhaps the other schedulers are given enough time to catch up as the best performing schedulers, such as Adagrad, starts slowing down the rate at which the MSE is decreased. However, as can be seen in HEATMAP FIG, Adagrad offers a 12\% \footnote{$\frac{constant MSE}{adagrad MSE} * 100 = \frac{0.018}{0.015} * 100 = 12\%$} decrease in MSE compared to the constant scheduler. It is clear from the results thus far that Adagrad is the best scheduler for approximating the Franke function when using linear regression with gradient descent. We believe that the reason Adagrad wins out is that these result are based on training for 100 epochs. The downfall of Adagrad in the literature seems to be that the aggressive decaying of the learning rate hinders a quick convergence deeper into training SOURCE https://towardsdatascience.com/learning-parameters-part-5-65a2f3583f7d. We are thus cautionary to claim that Adagrad will do the best job over a much larger amount of epochs when compared to adaptive schedulers that claim to improve this aspect.

To test this, we compared Adagrad with Adam over 5000 epochs. Both schedulers performed equally, both resulting in an MSE of 0.017 with Adagrad taking the edge with an MSE 0.0002 lower. Scikit obtained a test data MSE of 0.005, but followed the same trend in performance of dipping early and stabilizing after just a 100 epochs.
